### 2015-11-04 figures/Figure2-flowDiagram.tex
Added --ignore option. Iterated results.

\documentclass{article}

\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

%%%<
\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{5pt}%
%%%>

\begin{comment}
:Title: Simple flow chart
:Tags: Diagrams

With PGF/TikZ you can draw flow charts with relative ease. This flow chart from [1]_
outlines an algorithm for identifying the parameters of an autonomous underwater vehicle model.

Note that relative node
placement has been used to avoid placing nodes explicitly. This feature was
introduced in PGF/TikZ >= 1.09.

.. [1] Bossley, K.; Brown, M. & Harris, C. Neurofuzzy identification of an autonomous underwater vehicle `International Journal of Systems Science`, 1999, 30, 901-913

\end{comment}

\begin{document}
% Define the layers to draw the diagram
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\pagestyle{empty}
\renewcommand{\arraystretch}{0.75}

%POSITIVE TERMS:
%egrep 'befit|winner|encourag|metric|noisy|algorithms$|kernel$|methods$|procedure$|programs$|software$|technologies$|accuracy$|analysis$|benchmark|comparing$|comparison$|comprehensive$|estimation$|evaluation$|effectiveness$|metrics$|efficiency$|performance$|perspective$|quality$|rated$|suitable$|suitability$|superior$|survey$|cputime$|speed$|time$|duration$|fast$|slow$' wordScores.tsv | sort -nr | egrep -v 'mrsfast|mrfast|alignmentmethods' | head -n 5  | perl -lane 'print "{\\tiny $F[0]} & {\\tiny $F[1]} & {\\tiny $F[2]} & {\\tiny $F[3]}\\\\"'

%NEGATIVE TERMS:
%egrep 'mirnas$|associated$|disease$|genes$|human$|proteins$|genetic$|pathways$|mrna$|cdna$|functional$|clinical$|regulation$|predicted$|interaction$|review$|network' 
%cat wordScores.tsv | sort -nr | egrep -v 'xxxxxx' | tail -n 5  | perl -lane 'print "{\\tiny $F[0]} & {\\tiny $F[1]} & {\\tiny $F[2]} & {\\tiny $F[3]}\\\\"'

% Define block styles
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=5em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, line width=1mm, -latex']
\tikzstyle{cloud} = [draw, rectangle,fill=red!20, 
       text width=14em, text centered, rounded corners, minimum height=4em]
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]

\begin{tikzpicture}[->,node distance = 6cm, auto,thick,main node/.style={font=\sffamily\Large\bfseries}]
    % Place nodes
    \node [align=center]                              (dummy)  {}; 
%    \node [block]                  (dummy)  {}; 
    \node [block, left of=dummy,   node distance=4.5cm] (init1)  {Training articles};
    \node [block, right of=dummy,  node distance=4.5cm] (init2)  {Background articles};
    \node [cloud, below of=dummy,  node distance=1cm] (remove) {Remove high freq. words};
    \node [cloud, below of=remove, node distance=2.5cm] (freqs)  {Compute word \& di-word freqs};
    \node [cloud, below of=freqs,  node distance=4cm] (lods)   {Compute word scores: $s(word)=\log_2\left(\frac{f_{training}(word)+\delta}{f_{background}(word)+\delta}\right)$\\
\begin{tabular}{llll}
{\tiny logOdds} & {\tiny tnFreq} & {\tiny bgFreq} & {\tiny word}\\
{\tiny 5.28}    & {\tiny 0.0019} & {\tiny 0.0000} & {\tiny benchmarking}\\
{\tiny 5.21}    & {\tiny 0.0061} & {\tiny 0.0002} & {\tiny benchmark}\\
{\tiny 4.91}    & {\tiny 0.0011} & {\tiny 0.0000} & {\tiny noisy}\\
{\tiny 4.85}    & {\tiny 0.0022} & {\tiny 0.0001} & {\tiny metrics}\\
{\tiny 4.85}    & {\tiny 0.0003} & {\tiny 0.0000} & {\tiny encouragingly}\\
...\\
{\tiny -7.90}   & {\tiny 0.0000} & {\tiny 0.0024} & {\tiny disease}\\
{\tiny -8.02}   & {\tiny 0.0000} & {\tiny 0.0026} & {\tiny associated}\\
{\tiny -8.09}   & {\tiny 0.0000} & {\tiny 0.0027} & {\tiny mirnas}\\
\end{tabular}
};
    \node [cloud, below of=lods, node distance=4cm]     (score)  {Score \& rank candidate articles: $\sum_{word\in article} s(word)$};
    \node [block, left  of=score, node distance=4.5cm]    (cand)   {Candidate articles};
    \node [decision, below of=score, node distance=3.5cm] (decide) {Manually evaluate high scoring articles};

    % Draw edges
    \path [line] (init1)  -- (remove);
    \path [line] (init2)  -- (remove);
    \path [line] (remove) -- (freqs);
    \path [line] (freqs)  -- (lods);
    \path [line] (lods)  -- (score);
    \path [line] (cand)  -- (score);
    \path [line] (score)  -- (decide);
    \path (decide) edge[bend right=90, line width=1mm] node[anchor=init2,right]{no }(init2);
    \path (decide) edge[bend  left=90, line width=1mm] node[anchor=init1,right]{yes}(init1);

%    \path (remove.north) +(0,-2.3) node (clouds) {Build model};
    \path (remove.north) +(-1.5,-2.0) node (clouds) {Build model};

    \begin{pgfonlayer}{background}
        \path (remove.west |- remove.north)+(-0.5,0.3) node (a) {};
        \path (lods.south -| lods.east)+(+0.5,-0.3)    node (b) {};
        \path (clouds.east |- clouds.east)+(+3.5,-7.8)   node (c) {};

        \path[fill=yellow!20,rounded corners, draw=black!50, dashed]
            (a) rectangle (c);           
        \path (remove.north west)+(-0.2,0.2) node (a) {};

    \end{pgfonlayer}

\end{tikzpicture}

\end{document}