### 2015-12-17 figures/Figure2-flowDiagram.tex
Did a lot of stuff. Updated analysis. Regularised IFs.

\documentclass{article}

\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

%%%<
\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{5pt}%
%%%>

\begin{comment}
:Title: Simple flow chart
:Tags: Diagrams

With PGF/TikZ you can draw flow charts with relative ease. This flow chart from [1]_
outlines an algorithm for identifying the parameters of an autonomous underwater vehicle model.

Note that relative node
placement has been used to avoid placing nodes explicitly. This feature was
introduced in PGF/TikZ >= 1.09.

.. [1] Bossley, K.; Brown, M. & Harris, C. Neurofuzzy identification of an autonomous underwater vehicle `International Journal of Systems Science`, 1999, 30, 901-913

\end{comment}

\begin{document}
% Define the layers to draw the diagram
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\pagestyle{empty}
\renewcommand{\arraystretch}{0.75}

%POSITIVE TERMS:
%egrep 'befit|winner|encourag|metric|noisy|algorithms$|kernel$|methods$|procedure$|programs$|software$|technologies$|accuracy$|analysis$|benchmark|comparing$|comparison$|comprehensive$|estimation$|evaluation$|effectiveness$|metrics$|efficiency$|performance$|perspective$|quality$|rated$|suitable$|suitability$|superior$|survey$|cputime$|speed$|time$|duration$|fast$|slow$' wordScores.tsv | sort -nr | egrep -v 'mrsfast|mrfast|alignmentmethods' | head -n 5  | perl -lane 'print "{\\tiny $F[0]} & {\\tiny $F[1]} & {\\tiny $F[2]} & {\\tiny $F[3]}\\\\"'

%NEGATIVE TERMS:
%egrep 'mirnas$|associated$|disease$|genes$|human$|proteins$|genetic$|pathways$|mrna$|cdna$|functional$|clinical$|regulation$|predicted$|interaction$|review$|network' 
%cat wordScores.tsv | sort -nr | egrep -v 'xxxxxx' | tail -n 5  | perl -lane 'print "{\\tiny $F[0]} & {\\tiny $F[1]} & {\\tiny $F[2]} & {\\tiny $F[3]}\\\\"'

% Define block styles
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=12em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, line width=1mm, -latex']
\tikzstyle{cloud} = [draw, rectangle,fill=red!20, 
       text width=32em, text centered, rounded corners, minimum height=4em]
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=9em, text badly centered, node distance=3cm, inner sep=0pt]

\begin{tikzpicture}[->,node distance = 20cm, auto,thick,main node/.style={font=\sffamily\Large\bfseries}]
    % Place nodes
    \node [align=center]                              (dummy)  {}; 
%    \node [block]                  (dummy)  {}; 
    \node [block, left of=dummy,   node distance=9cm] (init1)  {\Huge Training articles};
    \node [block, right of=dummy,  node distance=9cm] (init2)  {\Huge Background articles};
    \node [cloud, below of=dummy,  node distance=2cm] (remove)   {\Huge Remove high freq. words};
    \node [cloud, below of=remove, node distance=3cm] (freqs)  {\Huge Compute word \& di-word freqs};
    \node [cloud, below of=freqs,  node distance=8cm] (lods)     {\Huge Compute word scores: $lo(word)=\log_2\left(\frac{f_{training}(word)+\delta}{f_{background}(word)+\delta}\right)$\\
\begin{tabular}{llll}
{\normalsize logOdds} & {\normalsize tnFreq} & {\normalsize bgFreq} & {\normalsize word}\\
{\normalsize 5.28}    & {\normalsize 0.0019} & {\normalsize 0.0000} & {\normalsize benchmarking}\\
{\normalsize 5.21}    & {\normalsize 0.0061} & {\normalsize 0.0002} & {\normalsize benchmark}\\
{\normalsize 4.91}    & {\normalsize 0.0011} & {\normalsize 0.0000} & {\normalsize noisy}\\
{\normalsize 4.85}    & {\normalsize 0.0022} & {\normalsize 0.0001} & {\normalsize metrics}\\
{\normalsize 4.85}    & {\normalsize 0.0003} & {\normalsize 0.0000} & {\normalsize encouragingly}\\
...\\
{\normalsize -7.90}   & {\normalsize 0.0000} & {\normalsize 0.0024} & {\normalsize disease}\\
{\normalsize -8.02}   & {\normalsize 0.0000} & {\normalsize 0.0026} & {\normalsize associated}\\
{\normalsize -8.09}   & {\normalsize 0.0000} & {\normalsize 0.0027} & {\normalsize mirnas}\\
\end{tabular}
};
    \node [cloud, below of=lods, node distance=8cm]     (score)  {\Huge Score \& rank candidate articles: $\sum_{i} lo(w_i)$};
    \node [block, left  of=score, node distance=10cm]    (cand)   {\Huge Candidate articles};
    \node [decision, below of=score, node distance=7cm] (decide) {\Huge Manually evaluate high scoring articles};

    % Draw edges
    \path [line] (init1)  -- (remove);
    \path [line] (init2)  -- (remove);
    \path [line] (remove) -- (freqs);
    \path [line] (freqs)  -- (lods);
    \path [line] (lods)  -- (score);
    \path [line] (cand)  -- (score);
    \path [line] (score)  -- (decide);
    \path (decide) edge[bend right=90, line width=1mm] node[anchor=init2,right]{\Huge no }(init2);
    \path (decide) edge[bend  left=90, line width=1mm] node[anchor=init1,right]{\Huge yes}(init1);

    \path (remove.north) +(7,-2.0) node[rotate=-90] (clouds) {\Huge Build model};

    \begin{pgfonlayer}{background}
        \path (remove.west |- remove.north)+(-1,0.3) node (a) {};
        \path (lods.south -| lods.east)+(-0.5,-0.3)    node (b) {};
        \path (clouds.east |- clouds.east)+(+1.75,-17.8)   node (c) {};

        \path[fill=yellow!20,rounded corners, draw=black!50, dashed]
            (a) rectangle (c);           
        \path (remove.north west)+(-0.2,0.2) node (a) {};

    \end{pgfonlayer}

\end{tikzpicture}

\end{document}